Reconstructing the Lensing Mass in the Universe \\ from Photometric Catalogue Data

### 2012-12-04 doc/paper-lcr.tex
edited Tom's second draft

\author[Collett \etal]{%
  Thomas~E.~Collett$^{1}$\thanks{\collettemail},
  Philip~J.~Marshall$^{2}$,
  Matthew~W.~Auger$^{1}$,
  Stefan~Hilbert$^{3}$,
\newauthor{%
  Sherry~H.~Suyu$^{4}$,
  Zachary~Greene$^{4}$,
  Tommaso~Treu$^{4}$\thanks{\packard},}
\newauthor{%
  Christopher~D.~Fassnacht$^{5}$,
  L\'eon~V.~E.~Koopmans$^{6}$,
  Roger~D.~Blandford$^{3}$} 
  \medskip\\
  $^1$\ioa\\
  $^2$\oxford\\
  $^3$\kipac\\
  $^4$\ucsb\\
  $^5$\davis\\
  $^6$\kapteyn
}

@@

High precision cosmological distance measurements towards individual objects
such as time delay gravitational lenses or type Ia supernovae are affected by
weak lensing perturbations by galaxies and groups along the line of sight.
In time delay gravitational lenses,
``external convergence,'' $\kappax$, can dominate the uncertainty in the inferred distances.
In this paper we attempt to reconstruct $\kappax$,
due to line of sight structure using a simple halo model.
 We use mock catalogues from the Millennium Simulation, and compare
our reconstructed $\pr(\kappax|\data)$ to ray-traced $\kappax$ truth values: 
taking into account realistic uncertainties on redshift and stellar masses, we
typically find that the reconstruction yields a factor of 2 improvement in precision of
over the global $\pr(\kappax)$. We find that the lowest $\kappax$ lines of sight have the best constrained 
$\pr(\kappax)$ after applying our reconstruction. We find that selecting systems by the precision with which
$\kappax$ can be estimated leads to a sample of unbiased
time delay distance estimates each with 1\% uncertainties due to external convergence effects.
Photometric data alone is sufficient to pre-select the best lenses and can be done
before investment in lightcurve monitoring. Preferentially selecting lines of
sight with high external shear (as might happen in an assembled sample of
quadruply-imaged time delay lenses) produces time-delay distance estimates with
typical external convergence uncertainties of 1.8\% but with systematic errors of 0.8\%.
We suggest areas for the improvement of this general analysis framework that
should allow yet more accurate cosmological inferences to be made.
\end{abstract}

@@

Every distant object we observe has had its apparent shape distorted,
and size and total brightness magnified (or demagnified) by a compound
weak gravitational lens constructed from all the mass distributed
between us and it. As \citet{Vale+White2003} and \citet{HilbertEtal2007}
showed, there are no empty lines of sight through our universe. This
fact makes gravitational lensing a potentially important source of
systematic error for any estimate of luminosity (or distance); this 
issue has been 
raised for \eg type Ia supernovae by
\citet[][]{Holz+Wald1998,Holz+Linder2005}, for gamma ray bursts by
\citet[][]{Oguri+Takahashi2006,Wang+Dai2011},  and for high redshift
galaxies by \citet{BradacEtal2009}, among others.

Along lines of sight containing strong gravitational lenses, the
perturbative effects of line of sight mass structure have been found to
be particularly important, with foreground and background structures
having a significant effect on the inferred lensing cross-section
\citep[\eg][]{WongEtal2012} and distance ratios
\citep[][]{DalalEtal2005}. Indeed, \citet{SuyuEtal2010} found that in
time delay lens cosmography the so-called ``external convergence''
$\kappax$ due to mass structures along the (unusually over-dense) line
of sight to the quadruply-imaged radio source B1608$+$656 was the
dominant source of uncertainty in their 5\% measurement of the time delay distance. 
Whilst it is rare for three galaxies to line up well enough for both of
the background sources to be strongly lensed \citep{GavazziEtal2008,CollettEtal2012a}, 
the large size of dark matter halos makes partial alignments -- such that they
act as perturbing weak lenses --  a near certainty. The large scale of the
perturbers means the external lensing perturbations can be approximated
by a quadrupole lens characterised only by external convergence and shear.
The external shear, $\gammax$ can potentially be recovered in the
mass-modelling of the lens, but the convergence is undetectable from the
image positions, shapes and relative fluxes; this is the famous {\emph{ 
mass-sheet degeneracy}} \citep[see e.g.][for details]{FalcoEtal1985}.

Large samples
of time-delay lenses are expected to be discovered in the next decade in
ground-based optical imaging surveys \citep{Oguri+Marshall2010}: both
the detectability of these lenses, and the selection of sub-samples of
them for further observation, could induce a selection function in external
convergence. External convergence must be included in analysis of time-delay
lenses else the inferred time-delay distance and cosmology be systematically biased.

Since the mass-sheet degeneracy prevents any estimate of $\kappax$ from the 
strong lens modelling additional information is required. Weak lensing measurements
can been used for rich lines of sight \citep{NakajimaEtal2009, FadelyEtal2009}, but 
for lines of sight with low galaxy density we must attempt to reconstruct the distribution
of mass along the line of sight. Attempts to reconstruct the mass distribution in strong
lens fields have focused on understanding the external shear which seems to be large in most strong lenses.
Surveys by \citet{Fassnacht+Lubin2002,AugerEtal2007,WilliamsEtal2006,MomchevaEtal2006}
all found groups of galaxies hosting or near to known gravitational lenses.
\citet{WongEtal2011} estimated $\Pr(\gammax)$ given the data of \citet{WilliamsEtal2006}
and \citet{MomchevaEtal2006} but found significant discrepancy between the predicted shear
distribution and the external shear demanded by the strong lens model. 
\citeauthor{WongEtal2011} also found that both line of
sight structures and the group of galaxies in each lens plane contribute
significant proportions of the shear.

For non-lens lines of sight magnification is a more useful quantity than convergence; \citet{GunnarssonEtal2006} used a galaxy halo model combined with simple scaling relations to reconstruct supernovae
lines of sight and found that the dispersion in
apparent source brightness due to lensing magnification could be reduced
by a factor of two. Conversely \citet{KarpenkaEtal2012} used the brightness dispersion in type Ia supernovae to infer the parameters of both their galaxy halo model and mass-to-light scaling relation. Both \citeauthor{KarpenkaEtal2012} and \citet{JonssonEtal2010} were able to 
detect lensing effects in the supernova data this way.

@@

The \MS results have been used to analyse real observations as well. 
\citet{SuyuEtal2010} selected \MS lines of sight by their apparent
galaxy over-density, in a 45 arcsec radius aperture down to $i < 24.5$,
to match the observed over-density towards the time delay lens
B1608$+$656 \citep{FassnachtEtal2011}.  The resulting distribution of
$\kappax$ values from the ray-tracing was taken to be an estimate of 
$\Pr(\kappax)$, which was then marginalised over when inferring the time
delay distance in this system. A similar procedure was used in \citet{SuyuEtal2012}.
In a companion paper to this one, Greene et al. (submitted) investigate
improvements to this method by weighting the galaxy counts by galaxy
mass, and perpendicular distance to the line of sight, again using the
\MS mock catalogues and their associated $\kappax$ values to construct
$\Pr(\kappax|\data)$. For the most over-dense lines of sight the
Green et al. results  show a $\sim$20\% improvement over
number counts alone, but little improvement for less dense lines.

In this paper, we combine the halo model reconstruction approach of
\citeauthor{GunnarssonEtal2006}, \citeauthor{JonssonEtal2010}, 
\citeauthor{WongEtal2011} and others, with the
idea of calibrating to simulations from \citeauthor{SuyuEtal2010}
Taking the $\kappax$ values from
\citeauthor{HilbertEtal2009}'s ray-tracing calculations as the ``truth''
that we have to recover, we use the \MS mock galaxy catalogues first to
{\it calibrate} the reconstruction, and account for unseen mass and
voids, and then to {\it test} the accuracy of the line of sight mass
reconstruction under this assumption. Probabilistically assigning mass
and redshift to every observed galaxy in a given observed field, we
generate Monte Carlo sample line of sight mass distributions, and so
construct the PDF $\Pr(\kappax|\data)$ for that field. We then emulate
the combination of many such PDFs to quantify the residual biases that
would be translated to the global (including cosmological) parameters.
In doing so, we aim to answer the following questions:

@@

\item Faint galaxies, filaments and dark structures will not appear in
any photometric object catalogue, but they will contribute convergence at
some level. How much of the total convergence comes from visible
galaxies, and how much effect do dark structures and voids have?

\item Can the true convergence be recovered from a calibrated halo model
reconstruction? What scatter and residual bias are induced by the
reconstruction process, and how might these be reduced in the future?

\item We have some choices to make when planning follow-up observations
of interesting sight-lines. How should we select objects to achieve the
highest accuracy convergence reconstruction? Can such a selection be
made robustly, using the reconstruction results?

@@

Throughout this paper magnitudes are given in the AB system and
we adopt the Millennium Simulation's ``concordance'' parameters for our reference cosmology, \ie
$h=0.73$, $\Omega_m=0.25$ and $\Omega_\Lambda=0.75$, where the symbols indicate
the Hubble Constant in units of 100 km s$^{-1}$ Mpc$^{-1}$ and the matter and
dark energy density of the Universe in units of the critical density.
.%\citep[e.g.\ ][]{KomEtal09}.

@@

\comments{In order to propagate the effect of line-of-sight perturbations into 
uncertainties on time-delay distances we require the probability density function (PDF) for the external convergence
given our available data $\data$, $\Pr(\xkappa|\data)$}
% PHIL's REVIEWS OF THE LITERATURE
\comments{\citet{GunnarssonEtal2006} investigated a related PDF,  for the
magnification, using a galaxy halo model. Applying empirical galaxy
scaling relations both to simulate mock catalogues and then to reconstruct
the line of sight mass distribution. While they did not explore highly
over-dense lines of sight, or the impact of groups and clusters, they
found that under their simplifying assumptions, the dispersion in
apparent source brightness due to lensing magnification could be reduced
by a factor of two.
%
\citet{WongEtal2011} estimated the PDF for the external shear,
$\Pr(\gammax|\data)$ from their survey data, running similar Monte Carlo
reconstructions of the mass in nine lens fields based on
\citeauthor{WilliamsEtal2006} and \citeauthor{MomchevaEtal2006}'s
photometric and spectroscopic measurements of galaxies close to the line
of sight. They then  compared the resulting predicted shear
distributions with the external shear demanded by the strong lens model.
Their reconstructions showed most of the shear to have been generated by
bright galaxies within 2~arcminutes of the lens, and that both line of
sight structures and the group of galaxies in each lens plane contribute
significant proportions of the shear. They found significant
discrepancies between the lens model shear and the Monte Carlo
predictions, but were unable to distinguish between the environment
modelling and the strong lens modelling as the cause of these
discrepancies.
%
\citet{KarpenkaEtal2012} used a halo model to predict deterministically 
the magnification due to
galaxies along the line of sight to type Ia supernovae, and inferred the
parameters of scaling relations between light and mass in these galaxies
from hundreds of supernova lightcone simultaneously. Following
\citet{JonssonEtal2010}, they subtract off the mean convergence to ensure that
the universal mass budget is balanced, and compare models where the halos are
all have either truncated singular isothermal profiles, or 
NFW profiles. Both groups were able to 
detect lensing effects in the supernova data this way.
}

\begin{figure*}
\includegraphics[width=\textwidth]{figs/viewofalightcone.eps}
\caption[magcut]{Four different views of a slightly over-dense \MS line
of sight, with a $\kappax$ of 0.03. 
{\it Top row, left:} The positions of galaxies projected on the sky. The area
of the circles is proportional to observed $i$-band flux; the brightest
object shown has an $i$ magnitude of 17.7. 
{\it Top row, centre:} The angular sizes of halos projected on the sky.
Red and blue regions lie within the NFW scale radius and virial radius
of each halo, respectively. There are essentially no empty lightcones. 
{\it Top row, right:} The individual $\kappax$ contributions of
each halo, assuming a \citet{BMO} profile and the \citet{Neto2007}
mass-concentration relation. Comparison of
this panel and the centre panel reveals the relative importance of
proximity to the line of sight.
{\it Bottom:} A view along the redshift axis, showing only halos with
$|x|<0.3$ arcmin. The area of the points is proportional to each halo's
mass: the most massive halo shown has $1.6\times10^{12}\Msun$. The
optical axis is shown by the dashed line, while the dotted lines mark
the lens and source planes for a B1608-like strong lens.}
\label{fig:lightcone}
\end{figure*}

@@

\section{Theoretical Background}
\label{sec:theory}
\comments{
Convergence, or Ricci focusing, occurs when a gravitational lens focuses
the rays in a given bundle. This focusing can cause distant objects to
appear brighter, and larger than they would if the lens were removed.
The convergence from an isolated mass sheet is the ratio of the projected surface
mass density ($\Sigma$) \devided by the
critical surface mass density at the redshift of the mass sheet
($\Sigma_{\rm cr}$),
\be
\kappax= \frac {\Sigma}
                              {\Sigma_{\rm cr}(\zd,\zs)}
\ee
where 
\be 
\label{eq:sigcrit} 
\Sigma_{\rm cr}(\zd,\zs) \equiv \frac{c^2 D_{\rm os}}{4 \pi G D_{\rm od} D_{\rm ds}}.
\ee
and the $D$'s are the angular diameter distances between the objects
referred to in the subscripts: o is the observer, d is the deflector and
s is the source. If the surface mass density of the sheet exceeds
$\Sigma_{\rm cr}$, multiple images of the source can occur, otherwise
only one image will be observed, although that image will still be
perturbed relative to the unlensed case.

@@

In strong lens systems where the source is time variable, the images do not 
vary simultaneously; the optical path length for each image is different due
to relativistic and geometric effects. The difference in optical path length
causes a time-delay between the light curves of each image. \citeauthor{FalcoEtal1985}
showed that the presence of additional matter (in the form of a mass-sheet)
along the line of sight has no observable effect except to rescale the time-delays.
As such, it is necessary to include $\kappax$ in the lens modelling, if
cosmological parameters are to be estimated accurately and precisely from
observed time delays \citep{SuyuEtal2010}.

@@

Mass distributions that are
physically associated with the lens galaxy will affect the stellar dynamics of
the lens galaxy, so that dynamical observations can break the internal
mass-sheet degeneracy by providing an additional estimate of the lens galaxy's
mass \citep[e.g.,][]{Koopmans2004,SuyuEtal2010}. In this work our aim is to quantify the
lensing effect of massive objects, regardless of their position along the line
of sight.

@@

@@

We construct mock photometric galaxy catalogues by selecting all \MS
galaxies within a circular aperture of a given angular radius centred on
the line of sight position, with apparent $i$-band magnitude brighter
than a given limit. The parent catalogue, from \citet{HilbertEtal2011},
contains galaxy positions and magnitudes, all of which include lensing
effects (deflections and magnifications). We also have access to the
underlying galaxy halo masses and redshifts, which we will use to
calibrate the reconstruction, and also to explore any sources of bias
and scatter. Since the true convergence is that of only the dark matter,
we focus on reconstructing the dark matter halos alone. Our model for
doing this is outlined in the next section.

@@

We require a method for estimating the
convergence at any point on the sky, given a catalogue of observed
galaxy properties such as positions, brightnesses, colours, and
possibly stellar masses and redshifts, for every galaxy in the field of
the lens, down to some magnitude limit. Since we cannot observe the
surface mass density of each halo directly, we need some way of
estimating the mass of each halo in the catalogue, so that we can compute
its contribution to the total $\kappax$ along the line of sight \citep[as
in \eg][]{GunnarssonEtal2006,WongEtal2011,KarpenkaEtal2012}. 
This mass assignment recipe
will be uncertain, but can be calibrated with cosmological simulations, which
represent a significant additional source of information to what is present in
the data.

Transforming an observed photometric galaxy catalogue into a 
catalogue of halo masses, positions, and redshifts, will enable us to
attempt a reconstruction of the convergence induced by every halo near a
line of sight. {We also need to account for the convergence due to dark structures and
the divergence due to voids; we do not model voids or dark structures, but their
effects are included statistically by calibration the the convergence in halos,
$\kappah$, against the ray-traced convergence, $\kappax$, for \MS lines of sight.}

@@

Cosmological dark matter simulations have shown that dark matter
halos are well-approximated by NFW profiles

\citep{NFW}, with a density
profile given by
\be\label{eq:rhonfw}
\rho(r) = 
\frac{\rho_0}{\left(r/r_{s}\right)
\left(1+r/r_{s}\right)^{2}},
\ee
where $r_{s}$ is a characteristic radius of the cluster, representing where
the density slope transitions from $r^{-1}$ to $r^{-3}$. This radius is related
to the virial radius of the halo by $r_{s}~=~r_{200}/c$, where $c$ is the concentration parameter, which can be estimated from the halo's mass,
using a mass--concentration relation. Typically more massive halos are less concentrated,
but there is some scatter; we use the relation of \citet{Neto2007} to estimate $c$
from the mass enclosed within $r_{200}$, which we denote as $M_{200}$. We find our results do not change if the \citet{MaccioEtal2008} Mass--concentration relation is used instead.

@@

Each halo in our catalogue then contributes to
the line of sight convergence by
\be
\label{eq:kappai}
\kappa_i =\Sigma_{i}/\Sigma_{\mathrm{cr}}(z_i,\zs),
\ee
where $\Sigma_{\mathrm{cr}}(z_i,\zs)$ is the critical surface density \comments{was defined in \Eref{eq:sigcrit}}.
Following the first order approximation of \citet{HilbertEtal2009}
outlined in \Sref{sec:MS} above, we compute 
the total convergence from all the halos along the line
of sight using
\be 
\label{eq:kappasummu}
\kappah = \sum_{i} \kappa_i.
\ee

>>>>

\author[Collett \etal]{%
  Thomas~E.~Collett$^{1}$\thanks{\collettemail},
  Philip~J.~Marshall$^{2}$,
  Matthew~W.~Auger$^{1}$,
  Stefan~Hilbert$^{3}$,
\newauthor{%
  Sherry~H.~Suyu$^{4}$,
  Zachary~Greene$^{4}$,
  Tommaso~Treu$^{4}$\thanks{\packard},
  Christopher~D.~Fassnacht$^{5}$,}
\newauthor{%
  L\'eon~V.~E.~Koopmans$^{6}$,
  Maru\v{s}a Brada\v{c}$^{5}$,
  Roger~D.~Blandford$^{3}$} 
  \medskip\\
  $^1$\ioa\\
  $^2$\oxford\\
  $^3$\kipac\\
  $^4$\ucsb\\
  $^5$\davis\\
  $^6$\kapteyn
}

@@

High precision cosmological distance measurements towards individual objects
such as time delay gravitational lenses or type Ia supernovae are affected by
weak lensing perturbations by galaxies and groups along the line of sight. In
time delay gravitational lenses, ``external convergence,'' $\kappax$, can
dominate the uncertainty in the inferred distances. In this paper we attempt
to reconstruct $\kappax$, due to line of sight structure using a simple halo
model.
%
We use mock catalogues from the Millennium Simulation, and calibrate and 
compare our reconstructed $\pr(\kappax|\data)$ to ray-traced $\kappax$ truth
values:  taking into account realistic uncertainties on redshift and stellar
masses, \phil{we find that the reconstruction provides an improvement in
precision of 50\% over the global $\pr(\kappax)$.} 
%
We find that the lowest $\kappax$ lines of sight have the best constrained 
$\pr(\kappax)$. We find that selecting the third of the systems with the
highest precision $\kappax$ estimates gives a  sub-sample of unbiased time
delay distance measurements \phil{with (on average) just} 1\% uncertainty due
to line of sight external convergence effects. Photometric data alone is
sufficient to pre-select the best-constrained lines of sight, and can be done
before investment in lightcurve monitoring. 
%
\phil{Conversely, we show that somehow preferentially selecting lines of sight
with high external shear could, with the reconstruction model presented here,
induce biases of up to 1\% in time delay distance.}  We suggest areas for the
improvement of this general analysis framework (including more sophisticated
treatment of high mass structures) that should allow yet more accurate
cosmological inferences to be made.

\end{abstract}

@@

Every distant object we observe has had its apparent shape distorted,
and size and total brightness magnified (or demagnified) by a compound
weak gravitational lens constructed from all the mass distributed
between us and it. As \citet{Vale+White2003} and \citet{HilbertEtal2007}
showed, there are no empty lines of sight through our universe. This
fact makes gravitational lensing a potentially important source of
systematic error for any estimate of luminosity (or distance); this 
issue has been 
raised for \eg type Ia supernovae by
\citet[][]{Holz+Wald1998,Holz+Linder2005}, for gamma ray bursts by
\citet[][]{Oguri+Takahashi2006,Wang+Dai2011},  and for high redshift
galaxies by \citet{WyitheEtal2011}, among others.

Along lines of sight containing strong gravitational lenses, the perturbative
effects of line of sight mass structure have been found to be particularly
important, with foreground and background structures having a significant
effect on the inferred lensing cross-section \citep[\eg][]{WongEtal2012} and
distance ratios \citep[][]{DalalEtal2005}. Indeed, \citet{SuyuEtal2010} found
that in time delay lens cosmography the so-called ``external convergence''
$\kappax$ due to mass structures along the (unusually over-dense) line of
sight to the quadruply-imaged radio source B1608$+$656 had to be included in
their analysis, and was the dominant source of uncertainty in their 5\%
measurement of the time  delay distance.  Large samples of time-delay lenses
are expected to be discovered in the next decade in ground-based optical
imaging surveys \citep{Oguri+Marshall2010};  the external convergence will
have to be understood increasingly well in order to prevent its correction
from dominating the systematic error budget.

While it is rare for three galaxies to line up well enough for both of
the background sources to be strongly lensed \citep{GavazziEtal2008,CollettEtal2012a}, 
the large size of dark matter halos makes partial alignments -- such that they
act as perturbing weak lenses --  a near certainty. The large scale of the
perturbers means the external lensing perturbations can be approximated
by a quadrupole lens characterised only by external convergence and shear.
The external shear, $\gammax$ can potentially be recovered in the
mass-modelling of the lens, but the convergence is undetectable from the
image positions, shapes and relative fluxes; this is the well-known {\emph{ 
mass-sheet degeneracy}} \citep[see e.g.][for details]{FalcoEtal1985}.

Since the mass-sheet degeneracy prevents any estimate of $\kappax$ from the 
strong lens modelling, additional information is required. Weak lensing measurements
can been used for rich lines of sight \citep{NakajimaEtal2009, FadelyEtal2009}, but 
for lines of sight with low galaxy density we must attempt to reconstruct the distribution
of mass along the line of sight. Attempts to reconstruct the mass distribution in strong
lens fields have focused on understanding the external shear which seems to be large in most strong lenses.
Surveys by \citet{Fassnacht+Lubin2002,AugerEtal2007,WilliamsEtal2006,MomchevaEtal2006}
all found groups of galaxies hosting or near to known gravitational lenses.
\citet{WongEtal2011} estimated $\Pr(\gammax)$ given the data of \citet{WilliamsEtal2006}
and \citet{MomchevaEtal2006} but found significant discrepancy between the predicted shear
distribution and the external shear demanded by the strong lens model. 
\citeauthor{WongEtal2011} also found that both line of
sight structures and the group of galaxies in each lens plane contribute
significant proportions of the shear.

For lines of sight without strong lenses, magnification is the more relevant
quantity than convergence; \citet{GunnarssonEtal2006} used a galaxy halo model
combined with simple scaling relations to reconstruct supernovae lines of
sight and found that the dispersion in apparent source brightness due to
lensing magnification could be reduced by a factor of two. Conversely
\citet{KarpenkaEtal2012} used the brightness dispersion in type Ia supernovae
to infer the parameters of both their galaxy halo model and mass-to-light
scaling relation. Both \citeauthor{KarpenkaEtal2012} and
\citet{JonssonEtal2010} were able to  detect lensing effects in the supernova
data this way.

@@

The \MS results have been used to analyse real observations as well. 
\citet{SuyuEtal2010} selected \MS lines of sight by their apparent
galaxy over-density, in a 45 arcsec radius aperture down to $i < 24.5$,
to match the observed over-density towards the time delay lens
B1608$+$656 \citep{FassnachtEtal2011}.  The resulting distribution of
$\kappax$ values from the ray-tracing was taken to be an estimate of 
$\Pr(\kappax)$, which was then marginalised over when inferring the time
delay distance in this system. A similar procedure was used in \citet{SuyuEtal2012}.
In a companion paper to this one, Greene et al. (submitted) investigate
improvements to this method by weighting the galaxy counts by galaxy
mass, and perpendicular distance to the line of sight, again using the
\MS mock catalogues and their associated $\kappax$ values to construct
$\Pr(\kappax|\data)$. For the most over-dense lines of sight their
results  show a $\sim$20\% improvement over
number counts alone, but little improvement for less dense ones.

In this paper, we combine the halo model reconstruction approach of
\citeauthor{GunnarssonEtal2006}, \citeauthor{JonssonEtal2010}, 
\citeauthor{WongEtal2011} and others, with the
idea of calibrating to simulations from \citeauthor{SuyuEtal2010}
Taking the $\kappax$ values from
\citeauthor{HilbertEtal2009}'s ray-tracing calculations as the ``truth''
that we have to recover, we use the \MS mock galaxy catalogues first to
{\it calibrate} the reconstruction, and account for unseen mass and
voids, and then to {\it test} the accuracy of the line of sight mass
reconstruction under this assumption. Probabilistically assigning mass
and redshift to every observed galaxy in a given observed field, we
generate Monte Carlo sample line of sight mass distributions, and so
construct the PDF $\Pr(\kappax|\data)$ for that field. We then emulate
the combination of many such PDFs to quantify the residual bias that
would be translated to the global (including cosmological) parameters, 
\phil{if left unaccounted for.}
In doing so, we aim to answer the following questions:

@@

\item Faint galaxies, filaments and dark structures will not appear in
any photometric object catalogue, but they will contribute convergence at
some level. How much of the total line of sight 
convergence comes from visible
galaxies, and how much effect do dark structures and voids have?

\item Can the true line of sight convergence be recovered from a calibrated
halo model reconstruction? What scatter and residual bias are induced by the
reconstruction process, and how might these be reduced in the future?

\item Both the detectability of lenses, and the selection of sub-samples of
them for further observation, could potentially induce a selection bias in
external convergence. How should we select objects to achieve the highest
accuracy convergence reconstruction? Can such a selection be made robustly,
using the reconstruction results?

@@

Throughout this paper magnitudes are given in the AB system and
we adopt the Millennium Simulation's ``concordance'' parameters for our reference cosmology, \ie
$h=0.73$, $\Omega_m=0.25$ and $\Omega_\Lambda=0.75$, where the symbols indicate
the Hubble Constant in units of 100 km s$^{-1}$ Mpc$^{-1}$ and the matter and
dark energy density of the Universe in units of the critical density.
%\citep[e.g.\ ][]{KomEtal09}.

@@

\comments{In order to propagate the effect of line-of-sight perturbations into 
uncertainties on time-delay distances we require the probability density function (PDF) for the external convergence
given our available data $\data$, $\Pr(\xkappa|\data)$}

% PHIL's REVIEWS OF THE LITERATURE
\comments{\citet{GunnarssonEtal2006} investigated a related PDF,  for the
magnification, using a galaxy halo model. Applying empirical galaxy
scaling relations both to simulate mock catalogues and then to reconstruct
the line of sight mass distribution. While they did not explore highly
over-dense lines of sight, or the impact of groups and clusters, they
found that under their simplifying assumptions, the dispersion in
apparent source brightness due to lensing magnification could be reduced
by a factor of two.}
%
\comments{\citet{WongEtal2011} estimated the PDF for the external shear,
$\Pr(\gammax|\data)$ from their survey data, running similar Monte Carlo
reconstructions of the mass in nine lens fields based on
\citeauthor{WilliamsEtal2006} and \citeauthor{MomchevaEtal2006}'s
photometric and spectroscopic measurements of galaxies close to the line
of sight. They then  compared the resulting predicted shear
distributions with the external shear demanded by the strong lens model.
Their reconstructions showed most of the shear to have been generated by
bright galaxies within 2~arcminutes of the lens, and that both line of
sight structures and the group of galaxies in each lens plane contribute
significant proportions of the shear. They found significant
discrepancies between the lens model shear and the Monte Carlo
predictions, but were unable to distinguish between the environment
modelling and the strong lens modelling as the cause of these
discrepancies.}
%
\comments{\citet{KarpenkaEtal2012} used a halo model to predict deterministically 
the magnification due to
galaxies along the line of sight to type Ia supernovae, and inferred the
parameters of scaling relations between light and mass in these galaxies
from hundreds of supernova lightcone simultaneously. Following
\citet{JonssonEtal2010}, they subtract off the mean convergence to ensure that
the universal mass budget is balanced, and compare models where the halos are
all have either truncated singular isothermal profiles, or 
NFW profiles. Both groups were able to 
detect lensing effects in the supernova data this way.}

\begin{figure*}
\includegraphics[width=\textwidth]{figs/viewofalightcone.eps}
\caption[magcut]{Four different views of a slightly over-dense \MS line
of sight, with a $\kappax$ of 0.03. 
{\it Top row, left:} The positions of galaxies projected on the sky. The area
of the circles is proportional to observed $i$-band flux; the brightest
object shown has an $i$ magnitude of 17.7. 
{\it Top row, centre:} The angular sizes of halos projected on the sky.
Red and blue regions lie within the NFW scale radius and virial radius
of each halo, respectively. There are essentially no empty lightcones. 
{\it Top row, right:} The individual $\kappax$ contributions of
each halo, assuming a \citet{BMO} truncated NFW 
profile and the \citet{Neto2007}
mass-concentration relation. Comparison of
this panel and the centre panel illustrates the relative importance of
proximity to the line of sight.
{\it Bottom:} A view along the redshift axis, showing only halos with
$|x|<0.3$ arcmin. The area of the points is proportional to each halo's
mass: the most massive halo shown has $1.6\times10^{12}\Msun$. The
optical axis is shown by the dashed line, while the dotted lines mark
the lens and source planes for a B1608-like strong lens.}
\label{fig:lightcone}
\end{figure*}

@@

\section{Theoretical Background}
\label{sec:theory}

\comments{
Convergence, or Ricci focusing, occurs when a gravitational lens focuses
the rays in a given bundle. This focusing can cause distant objects to
appear brighter, and larger than they would if the lens were removed.
The convergence from an isolated mass sheet is the ratio of the projected surface
mass density ($\Sigma$) \devided by the
critical surface mass density at the redshift of the mass sheet
($\Sigma_{\rm cr}$),
\be
\kappax= \frac {\Sigma}
                              {\Sigma_{\rm cr}(\zd,\zs)}
\ee
where 
\be 
\label{eq:sigcrit} 
\Sigma_{\rm cr}(\zd,\zs) \equiv \frac{c^2 D_{\rm os}}{4 \pi G D_{\rm od} D_{\rm ds}}.
\ee
and the $D$'s are the angular diameter distances between the objects
referred to in the subscripts: o is the observer, d is the deflector and
s is the source. If the surface mass density of the sheet exceeds
$\Sigma_{\rm cr}$, multiple images of the source can occur, otherwise
only one image will be observed, although that image will still be
perturbed relative to the unlensed case.

@@

In strong lens systems where the source is time variable, the images do not 
vary simultaneously; the optical path length for each image is different due
to relativistic and geometric effects. The difference in optical path length
causes a time-delay between the light curves of each image.
\citet{FalcoEtal1985} showed that the presence of additional matter (in the
form of a mass-sheet) along the line of sight has no observable effect except
to rescale the time-delays. As such, it is necessary to include $\kappax$ in
the lens modelling, if cosmological parameters are to be estimated accurately
and precisely from observed time delays \citep{SuyuEtal2010}.

@@

Mass distributions that are physically associated with the lens galaxy will
affect the stellar dynamics of the lens galaxy, so that dynamical observations
can break the internal mass-sheet degeneracy by providing an additional
estimate of the lens galaxy's mass
\citep[e.g.,][]{Koopmans2004,SuyuEtal2010}.  Our aim is to quantify the
lensing effect of massive objects, regardless of their position along the line
of sight. \phil{However, in this paper we restrict ourselves to reconstructing
the line of sight portion of the mass distribution in any given field, which
allows us to work in the weak lensing regime. Treatment of sightlines
containing strong lensing events, and the impact of physically-associated mass
distributions on both the strong lensing and stellar dynamical effects are
postponed to further work.}

@@

\comment{PJM: This means our sightlines are not typical of strong lenses, because they
don't include the local overdense region typically found near massive lens
galaxies. Our lines of sight have typical mass distributions for a lens line
of sight {\it after excision of the lens plane}. I have tried to explain
this above.}

@@

We construct mock photometric galaxy catalogues by selecting all \MS galaxies
within a circular aperture of a given angular radius centred on each randomly
selected line of sight position, with apparent $i$-band magnitude brighter
than a given limit. The parent catalogue, from \citet{HilbertEtal2011},
contains galaxy positions and magnitudes, all of which include lensing effects
(deflections and magnifications). We also have access to the underlying galaxy
halo masses and redshifts, which we will use to calibrate the reconstruction,
and also to explore any sources of bias and scatter. Since the true
convergence is that of only the dark matter, we focus on reconstructing the
dark matter halos alone. Our model for doing this is outlined in the next
section.

@@

We require a method for estimating the convergence at any point on the sky,
given a catalogue of observed galaxy properties such as positions,
brightnesses, colours, and possibly stellar masses and redshifts, for every
galaxy in the field of the lens, down to some magnitude limit. Since we cannot
observe the surface mass density of each halo directly, we need some way of
estimating the mass of each halo in the catalogue, so that we can compute its
contribution to the total $\kappax$ along the line of sight \citep[as in
\eg][]{GunnarssonEtal2006,WongEtal2011,KarpenkaEtal2012}.  This mass
assignment recipe will be uncertain, and also incomplete, but it can be
calibrated with cosmological simulations, which represent a significant
additional source of information to what is present in the data.

Transforming an observed photometric galaxy catalogue into a  catalogue of
halo masses, positions, and redshifts will enable us to attempt a
reconstruction of the convergence induced by every halo near a line of sight. 
We also need to account for the convergence due to dark structures and the
divergence due to voids; we do not model voids or dark structures, but their
effects are included statistically by calibrating the convergence in
halos, $\kappah$, against the ray-traced convergence, $\kappax$, for \MS lines
of sight.

@@

Cosmological dark matter simulations have shown that dark matter
halos are well-approximated by NFW profiles 
\citep{NFW}. We assume each halo to have a spherical mass distribution with 
density given by
\be\label{eq:rhonfw}
\rho(r) = 
\frac{\rho_0}{\left(r/r_{s}\right)
\left(1+r/r_{s}\right)^{2}}.
\ee
Here, $r_{s}$ is a characteristic scale 
radius of the cluster, representing the point where
the density slope transitions from $r^{-1}$ to $r^{-3}$. This radius is
related to the virial radius of the halo by $r_{s}~=~r_{200}/c$, where $c$ is
the concentration parameter, which can be estimated from the halo's mass,
using a mass--concentration relation. Typically more massive halos are less
concentrated, but there is some scatter; we use the relation of
\citet{Neto2007} to estimate $c$ from the mass enclosed within $r_{200}$,
which we denote as $M_{200}$. We find our results do not change if the
\citet{MaccioEtal2008} mass--concentration relation is used instead.

@@

Many studies have shown that galaxies have total mass density profiles that
are approximately isothermal in their inner regions
\citep[\eg][]{AugerEtal2010} due to the more concentrated stellar mass
component; however, at radii of tens of kpc (tens of arcsec), we expect the
halo density profile to be dark matter dominated, and fall off faster than
$r^{-2}$. This has been confirmed by galaxy-galaxy weak lensing measurements 
\citep[\eg][]{Mandelbaum,Leauthaud}. We note that the two halo term that
dominates the galaxy-galaxy lensing signal at large radii is explicitly taken
into account in our model, where every galaxy is assigned a dark matter halo.
